% These notes should be for one class period. 
\documentclass[12pt]{../notes}

% Command for Questions
%\question{}

% Command for Notes
% \note{}

% Code to create a minipage where you can type in class notes. 
%%\begin{minipage}[l][2cm][c]{\textwidth}
\begin{comment}

\end{comment}
%%\end{minipage}


% Begin Document
%==============================================================================
\begin{document}
% Include the Title of the Handout
\ntitle{Handout 1.2: Introduction to Hypothesis Testing}

\question{Determine whether or not the following statements are true or false:
\begin{itemize}
\item The p-value is the probability that the null hypothesis is true. %\note{(FALSE)}
\item We reject a null hypothesis when the p-value is small. %\note{(TRUE)}
\item If the p-value is very small, it is not possible that the null hypothesis is true. %\note{(FALSE)} 
\item The difference between the sample mean and the population mean is all that matters in the test statistic. %\note{(FALSE)}
\end{itemize}
}

\question{What is the different between a practical difference and a ``significant'' difference?}

\begin{minipage}[l][6cm][c]{\textwidth}
\begin{comment}
\note{A practical difference is one in which the magnitude of the difference is large enough to matter in the context of the problem. (Example: a new drug lowers cholesterol, on average by three points. Such a reduction is too small to justify taking the drug, regardless of the statistical significance.}

$\newline$
\note{A practical difference is not always significant, and a significant difference is not always practical.}

$\newline$
\note{When sample sizes are LARGE, nearly every difference is flagged as significant, even if the actual difference between groups is inconsequential.}
\end{comment}
\end{minipage}

\question{Can you think of an example when our primary motive for creating a model is to create accurate predictions? How about an example where the primary motive is determining the significance of the coefficients?}

\begin{minipage}[l][3cm][c]{\textwidth}
\begin{comment}
\note{Accuracy: Predicting the market value of a house given square footage, lot size, etc.}

\nspace
\note{Significance: Determining if there is a statistically significant gender bias in pay, after accounting for other demographic factors.}
\end{comment}
\end{minipage}


























% End the Document
%==============================================================================
\end{document}